### 2014-06-24 tex/motivation.tex
took a stab at some text - mainly the abstract and intro

\section{Background and Motivation}
\label{motivations}

Example