### 2014-11-09 tex/lazy.tex
added the new supplemenatal file

\section{Generalized Symbolic Execution}

Recall that in generalized symbolic execution (GSE) for lazy
initialization of symbolic references consists of three key points of
non-determinism where each symbolic reference can be initialized
non-deterministically to null, a new instance of the symbolic
reference, or aliases to symbolic references of the same type
previously initialized. The initialization in GSE consists of creating
branches in the execution tree for all the non-deterministic choices.

\input{initialize-lazy}
\input{lazy-fig}

\input{lazy-eq}