### 2014-11-12 tex/lazy.tex
Reworked supplemental upto definitions and proofs.

\section{Generalized Symbolic Execution}

The specialization of Javalite to generalized symbolic execution with
lazy initialization using the summary heap is in \figref{fig:lazyInit}
for the initialization and \figref{fig:lazy} for how the
initialization is used. Specifically, the field-access rule in
\figref{fig:lazy} uses initialization, $\rightarrow_I^*$, to be sure
the accessed field is instantiated. Initialization in generalized
symbolic execution never happens for more than one object on any use
of $\rightarrow_I^*$: $\Lambda$ is either empty or contains exactly one
location.

An uninitialized object may be lazily initialized
non-deterministically to null, a new instance of the object, or an to
an object of the same type also initialized lazily. The
non-deterministic initialization creates branches in the execution
tree during state space exploration.

The $\rgse$ relation is a union over branching points on control flow
to support the final proof of the existence of a
bisimulation. Although it is apparent that there is non-determinism in
initialization, it is not the initialization where control flow
diverges in program execution. Rather, that occurs at branches and
exceptions, and that is reflected in the composition of $\rgse$.

\input{initialize-lazy}
\input{lazy-fig}