### 2015-01-15 cav-2015-tex/motivation.tex

\paragraph{Why would you want to do lazy initialization, instead of using pre-generated input heaps?}
\begin{compactdesc}
\item[Generate a minimal set of independent test inputs:] 
If you generate a fixed set of 1000 data structures at the beginning,
you don’t know which of those 1000 data structures will actually
result in different execution paths. Since many inputs may result in
identical execution paths, the “path-independence” of those inputs is
uncertain. Consider the case of a linked list, where each node
contains two references, one for a data object, and one for the next
node in the list. Suppose you want to test a function that makes
structural transformations on linked lists. Perhaps your function
removes loops, counts the number of nodes, or pops the first node off
the list. If you tried every possible linked list as an input to your
function, you would find that many of your lists led to identical
execution paths, simply because the “data” reference was never
accessed. Without lazy initialization, you would have to know
beforehand which parts of your data structure will get accessed, and
which will not. With lazy initialization, we test only independent
sets of inputs, without any prior knowledge of the data structure. In
fact, we generate a minimal set of independent test inputs within the
given bounds.

\item[Make per-path search decisions:] 
With lazy initialization, we can make search decisions dynamically
based on the machine state. For example, suppose you want to limit the
number of nodes that any given execution path might attempt to
access. Note that this is very different than bounding the input heap
to a set size. Without lazy initialization, you would have to know
exactly which reference chains any given execution path will follow
beforehand. However, with lazy initialization, it is simple to keep
track of the number of nodes that have been accessed at run-time, and
limit the size of the lazy-initialized input heap dynamically. Some
examples of types of per-path bounding metrics which are possible
using lazy initialization include: length of reference chains, number
of cycles in the graph, number of nodes accessed by the program, the
amount of memory available to the test machine, total number of memory
reads/writes. (Note that some of these things may be difficult to
relate exactly back to the concrete heaps)

\item[Generate test inputs based on per-path execution properties:] 
Per-path search bounds can be combined with the test-input generation
capability to answer questions like “Which test inputs will generate
fewer than 20 memory read operations?”.

\item[Cope with certain types of unbounded inputs:] 
In some cases, lazy initialization can exactly model program behavior
even when the size of the input heap is unbounded. For example,
suppose we have a function that pushes a node on to the front of a
linked list. In this case, lazy initialization will correctly model
all program behaviors for linked lists of arbitrary size.

\item[Less work for the programmer:] 
You don’t need to generate any special data invariants specific to the
analysis. They are generated on-the-fly from invariants written in
normal program code.

\item[Delay choices about data structures:] 
If you are exploring a program with 1000 possible input structures, an
execution tree with 1000 leaves will have fewer states that a set of
1000 separate paths. Even in the case where all paths are merged into
a single path, the constraints on the earlier states will be simpler
if you don’t have to distinguish all 1000 possible inputs.
\end{compactdesc}

\paragraph{There are lots of symbolic heap representations. Why hasn’t anyone done lazy initialization before?}
\begin{compactitem}
\item[Inductively Defined Symbolic Locations:] 
Earlier techniques using similar representations (SATURN, the Dilligs)
have been limited by inductively defined symbolic locations. Suppose
you were to define symbolic locations as being either a method
parameter or a dereferenced symbolic location. This works well in
heaps without loops, because dereferenced symbolic location are
guaranteed to be “smaller” (closer to the root of the heap) than their
targets. However, this inductive definition breaks down in the case of
loops. Some recent work uses with non-recursive data structures,
(e.g. Koushik’s stuff), but they don’t do lazy initialization.

\item[Flat memory models:] 
Many prior techniques make no special treatment for object-oriented
memory models (Boogie, PEX, etc.). Typically the memory is modeled as
a flat address space instead of a more high-level heap. There are many
types of interactions that are possible in this memory model, such as
pointer arithmetic and array indexing, which are not possible in an
object-oriented system. Accounting for these interactions make lazy
initialization challenging to implement.

\item[Path splitting during memory access:] 
Several existing methods for doing lazy initialization during symbolic
execution introduce nondeterminism during memory access operations
(GSE, Lazier\#, etc.). These new points of nondeterminism contribute
to path explosion, rendering these methods infeasible for all but the
smallest heap fragments.
\end{compactitem}

\paragraph{What’s special about using a graph representation? Why not represent the heap as an equation?}
\begin{compactdesc}
\item[Fewer Solver Calls:] 
With a graph-based representation, we can answer structural questions
about the heap without consulting the solver. For example, when
comparing two references, if the intersection of their value sets is
empty, we can surmise that they are in all cases unequal. Some heap
properties that may be resolvable from structural properties alone
are: equivalence of references, connectivity of heap fragments, upper
bounds on pointer chain length, upper bounds on number of objects in
the heap, and the absence of loops / aliasing.

\item[Non-inductive definition of references:] 
A graph-based representation can provide a non-inductive definition of
the heap. Equation-based systems for modeling memories tend to
introduce lots of auxiliary variables, which can complicate
dereferencing and may place restrictions on heap morphology.  
\end{compactdesc}

\paragraph{What’s special about the “bipartite graph” representation?}
\begin{compactdesc}
\item[Object-oriented heap:] 
Our bipartite graph is a non-recursive model more suited to OO
programming than one based on pointers. In many OO programming
environments, you don’t have to worry about nasty things like pointer
arithmetic, so a cleaner model reduces wasted effort. You might say
the bipartite graph is how we’ve modeled a “theory of objects” for
heap constraint problems.
\end{compactdesc}