### 2015-04-20 ase-2015-tex/st-def.tex

\begin{definition}
\label{def:ST}
The function $\mathbb{ST}(\cfgnt{L},\cfgnt{r},\phi,\phi_g)$
strengthens every constraint from the reference $\cfgnt{r}$ with
$\phi$ and retaining location-constraint pairs that are satisfiable
after this strengthening with the inclusion of the global heap
constraint $\phi_g$:
\[
\begin{array}{rcl} 
\mathbb{ST}(\cfgnt{L},\cfgnt{r},\phi,\phi_g) & = & \{ (\phi\wedge\phi^\prime\ \cfgnt{l}^\prime) \mid (\phi^\prime\ \cfgnt{l}^\prime)\in \cfgnt{L}(\cfgnt{r})\wedge\mathbb{S}(\phi\wedge\phi^\prime\wedge\phi_g) \}
\end{array}
\]
\end{definition}