### 2015-09-10 vmcai-2016-tex/intro.tex
Re-worked parts of the introduction for clarity. Added transition material to sections on GSE and symbolic initialization.

\section{Introduction}

%In recent years symbolic execution has provided the basis for various
%software testing and analysis techniques, e.g., regression
%analysis~\cite{backes:2012,Godefroid:SAS11,Person:FSE08,person:pldi2011,Ramos:2011,Yang:ISSTA12},
%data structure repair~\cite{KhurshidETAL05RepairingStructurally},
%dynamic discovery of
%invariants~\cite{CsallnerETAL08DySy,Zhang:ISSTA14}, and
%debugging~\cite{Ma:2011}.
%
%Recent work has explored the strengths and limitations of
%tools which leverage constraint solvers, e.g., concolic execution,
%identifying symbolic dereferencing and updating operations as areas
%for potential improvement~\cite{Chen:2013,Qu:2011}.

In symbolic execution, the values of program variables are represented as constraints over the program's inputs. During the course of an execution, a program variable may assume a number of possible values under a variety of different conditions. One way to represent this behavior is by pairing each possible value with the constraints under which it is assumed. A set of such value-constraint pairs, called a \emph{guarded value set}\footnote{Guarded value sets are variously referred to in the literature as \emph{guarded location sets}, \emph{symbolic value sets}, or \emph{value summaries}. The term guarded value set is sometimes abbreviated in this text as~\emph{value set}.}, represents the state of a program variable.

Guarded value sets are rising in importance in the representation of reference values. First appearing in Verification-Condition Generator (VCG) style techniques~\cite{Xie:2005,Dillig:2011}, guarded value sets are a convenient way to represent sets of heap shapes. More recently, guarded value sets are being used in symbolic execution for the purposes of modularization~\cite{Yorsh:2008}, state merging~\cite{Sen:2014}, and for determining invariants~\cite{Ferrara:2014,Torlak:2014}.

A common hurdle to using guarded value sets is the treatment of symbolic heap inputs. For many applications, it is desirable to model the input heap as a black box, with operations over the input references providing the only means of divining its structure. Precisely modeling operations on black-box heap inputs is challenging because it requires formulating logical predicates over an input domain that contains a potentially unbounded number of hidden references~\cite{Chen:2013,Qu:2011}. Generalized symbolic execution (\gsetxt{}) methods~\cite{GSE03,Deng:2007} provide an accurate solution, but they quickly produce an overwhelming number of execution paths for all but the simplest heap shapes. Efforts to extend~\gsetxt{} principles to guarded value sets have failed to produce provably equivalent results. Thus, despite the advantages of using guarded value sets, it remains an open question how they can be used in symbolic execution to automatically model all possible program behaviors in the case of an arbitrary input heap.

The contribution of this work is to build on previous efforts, by demonstrating an initialization scheme that enables pointer dereferencing in an unknown symbolic input heap, and by proving that the initialization scheme is sound and complete with respect to properties provable by symbolic execution. A proof-of-concept implementation also demonstrates the efficiency of value-summary representations, as compared to existing case-splitting methods.