### 2015-09-12 vmcai-2016-tex/intro.tex
Revision and refactor

\section{Introduction}

%In recent years symbolic execution has provided the basis for various
%software testing and analysis techniques, e.g., regression
%analysis~\cite{backes:2012,Godefroid:SAS11,Person:FSE08,person:pldi2011,Ramos:2011,Yang:ISSTA12},
%data structure repair~\cite{KhurshidETAL05RepairingStructurally},
%dynamic discovery of
%invariants~\cite{CsallnerETAL08DySy,Zhang:ISSTA14}, and
%debugging~\cite{Ma:2011}.
%
%Recent work has explored the strengths and limitations of
%tools which leverage constraint solvers, e.g., concolic execution,
%identifying symbolic dereferencing and updating operations as areas
%for potential improvement~\cite{Chen:2013,Qu:2011}.

In symbolic execution, the values of program variables are represented as constraints over the program's inputs. During the course of an execution, a program variable may assume a number of possible values under a variety of different conditions. One way to represent this behavior is by pairing each possible value with the constraints under which it is assumed. A set of such value-constraint pairs, called a \emph{guarded value set}\footnote{Guarded value sets are variously referred to in the literature as \emph{guarded location sets}, \emph{symbolic value sets}, or \emph{value summaries}. The term guarded value set is sometimes abbreviated in this text as~\emph{value set}.}, represents the state of a program variable.

Guarded value sets are rising in importance in the representation of reference values. First appearing in Verification-Condition Generator (VCG) style techniques~\cite{Xie:2005,Dillig:2011}, guarded value sets are a convenient way to represent sets of heap shapes. More recently, guarded value sets are being used in symbolic execution for the purposes of modularization~\cite{Yorsh:2008}, state merging~\cite{Sen:2014}, and for determining invariants~\cite{Ferrara:2014,Torlak:2014}.

A common hurdle to using guarded value sets is the treatment of symbolic heap inputs. For many applications, it is desirable to model the input heap as unconstrained, with operations over the input references providing the only means of divining its structure. Precisely modeling operations on unconstrained heap inputs is challenging because it requires formulating logical predicates over an input domain that contains a potentially unbounded number of hidden references~\cite{Chen:2013,Qu:2011}. Generalized symbolic execution (\gsetxt{}) methods~\cite{GSE03,Deng:2007} provide an accurate solution, but they quickly produce an overwhelming number of execution paths for all but the simplest heap shapes. Efforts to extend~\gsetxt{} principles to guarded value sets have failed to produce provably equivalent results \cite{LISTEFFORTS}. Thus, despite the advantages of using guarded value sets, it remains an open question how they can be used in symbolic execution to automatically model all possible program behaviors in the case of an arbitrary input heap.

The contribution of this work is a symbolic initialization scheme for uninitialized pointers in a fully symbolic heap. Where GSE lazily instantiates uninitialized pointers to either reference NULL, a new instance of the correct type, or an alias to a previously initialized object, the symbolic initialization in this work creates a guarded value set expressing all of these eventualities in the same symbolic heap. Where GSE branches the search space on each choice, symbolic initialization does not. Symbolic initialization only branches on reference compares, partitioning the symbolic to indicate heaps where the references alias and heaps where the references do not.

This paper includes a proof that symbolic initialization is sound and complete with respect to properties provable by GSE. Such a results is important because it means that symbolic initialization can be used to create input heaps for other analyses in a way that is provably correct with regards to GSE: there are no missing heaps and there are no extra heaps.

A proof-of-concept implementation of symbolic initialization is reported in this paper. It demonstrates the efficiency of~\symtxt{}, as compared to existing~\gsetxt{} methods. The use of the guarded value sets does require extra interaction with an SMT solver, but the savings in representing multiple heap shapes in a single representation overcomes the cost of the solver calls. In fact, the comparison shows an exponential increase in the number of heaps that can be analyzed in a set of benchmark examples meaning that other approaches based on~\gsetxt\ fail to complete.