### 2015-10-28 vmcai-2016-tex/value-set-def.tex

\begin{definition}
\label{def:VS}
The $\mathbb{VS}$ function constructs the value set given a
heap, reference, and desired field.
\[
\begin{array}{rcl}
  \mathbb{VS}(L,R,\phi_g,\cfgnt{r},\cfgnt{f}) & \defeq &
  \{(\phi\wedge\phi^\prime\ l^\prime) \mid \exists
  l\ ((\phi\ l) \in L(\cfgnt{r})\ \wedge \\ & &
  \ \ (\phi^\prime\ l^\prime) \in L(R(l,\cfgnt{f}))\ \wedge \\ & &
  \ \ \mathbb{S}(\phi\wedge\phi^\prime\wedge \phi_g))\}
\end{array}
\]
\end{definition}